---
title: "Case Study B - Pain Relief Medication"
author: "Oumaima Al Qoh, Francisco Arrieta, Lucia Camenisch, Manuela Giansante, Emily Schmidt, Camille Beatrice Valera"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true # creating a table of contents (toc)
    toc_float: 
      collapsed: false # toc does not collapse and is shown as a sidebar (toc_float)
    number_sections: true # document sections are numbered
    theme: cosmo
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```

In the drug development process one of the important steps is to determine the minimum dosage amount of the drug that achieves efficacy. Sometimes this may involve estimating combinations of drugs that achieve maximum efficacy with a
minimum drug amount.

The experiment in this case study consists of administering a combination of two drugs to 10 mice at different doses. The drugs are morphine and marijuana and are used as pain relief medication (‘tranquilizers’). In order to measure their effect ‘flick tail’ tests were performed for two drugs and their combination.

The client’s objective is to detect whether a synergy exists between these two drugs. In addition, the client is also interested to know the effective dosage at which 50% of the subjects would be expected to respond.


```{r}
library(kableExtra)    # Tables
library(dplyr)         # Mutation for dummy variable
library(caret)         # Dummy variable
library(tidyr)         # Can be used to “gather” a key-value pair across multiple columns
library(cowplot)       # plot_grid() function
library(GGally)        # Allows to build a great scatterplot matrix
library(fastDummies)   # Create dummy variables
library(reshape2)      # melt() function
library(data.table)    # data.table format

#source("VIF.R")
```



```{r}
# Load your data, Bpain.csv
pain  = fread("Bpain.csv")
pain1 = fread("Bpain.csv")
```


**Lucia comment:** Why are we creating dummies? And I think that if we do create them, we shouldn't set them up in this way, but a dummy for A and a dummy for B since we have two drugs which are either present or not (no need for O and AB as columns)

```{r}
# Moved reps to the end of the dataframe
pain <- pain[, c(1:5, 7, 6)]

# Remove Observations to review quantitative values
pain <- pain[,-1] 

# Create dummy variables for Set
pain <- dummy_cols(pain, select_columns = "Set")

#Remove set since now dummies are available
pain <- pain[,-1] 
```

### **Exploratory Data Analysis**
-   obs: index number;

-   Set: one of four possible combinations (O = \'control\', A =
    \'morphine\', B= \'marijuana\', AB = \'combination of both drugs\');

-   Morphine: dose of morphine \'sulfate\' (mg/kg) injected into study
    mice
    --- the range is 0 to 8;

-   Marijuana: dose of marijuana \'Delta9-THC\' (mg/kg) injected into
    study
    mice --- the range is from 0 to 16;

-   Flick: the number of mice that did not flick the tail after being
    applied
    a heat stimulus from beneath ( \'flick tail\' test);

-   Reps: number of repetitions (with different mice);

-   Prob: proportion of mice (out of 10) that did not flick their tails
    after
    being administered a heat stimulus for a given drug combination (=
    Flick/Reps).
```{r, message=FALSE, warning=FALSE}
dim(pain) # Get or set the dimension of the specified matrix, array or data frame

kable(head(pain)) |>  kable_classic() |> 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive")) |> 
  kable_styling(position = "center") # Use kable to design table

str(pain) # Compactly displaying the internal structure of a R object

which(is.na(pain)) # Function returns the positions with missing values, 0 missing values
```

```{r}
# Produce result summaries
CSUM <- data.frame(mean = sapply(pain[,c(1:9)], mean,na.rm = T) 
                ,median = sapply(pain[,c(1:9)], median,na.rm = T)
                ,min = sapply(pain[,c(1:9)], min,na.rm = T)
                ,max = sapply(pain[,c(1:9)], max,na.rm = T)
                ,sd = sapply(pain[,c(1:9)], sd,na.rm = T))
colnames(CSUM) = c("Mean","Median","Min","Max","Standard Deviation")
kable(head(CSUM, 10)) %>% kable_classic() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive")) %>%
  kable_styling(position = "center") # Use kable to design table
```


**Lucia Comment:** dummies are not quantitative variables, should not be included in histograms. We should do a bar chart for the categorical variable sets instead. Variable Rep should also probably be excluded from the histograms (maybe from the dataset? as it brings no information, it is a set constant which would have its place in the metadata of our dataset rather than in the dataset itself)

```{r}
histograms <- ggplot(gather(data = pain),aes(value)) +
  geom_histogram(aes(y=..density..),bins = 10, color = "white", fill = "blue") + # Creates bin sizing and sets the lines as white
  geom_density(alpha= .2, fill="#56B4E9") +
  facet_wrap(~key,scales = "free") + # Converting the graphs into panels
  ggtitle("Quantitative Variable Analysis") + # Title name
  ylab("Count") + xlab("Value") + # Label names
  theme_classic() # A classic theme, with x and y axis lines and no grid lines

histograms
```


- the morphine group has more units
- the marijuana group has the least units
- the marijuana doses progress with different sized steps (maybe let's ask why?)
##### morphine
- between the dose 3 and 4 of morphine the same amount of subjects did not flick their tails (4/10)
- between the morphine dose 6 an d7 9/10 did not flick their tail
- at dose 8 (the highest) of morphine only not flick their tails
##### marijuana
- none flicked their tails after the first 2 doses (1-2) of marijuana;
- dose 4 - 8 of marijuana, 9/10 did not flick
- dose 16 , none moved
##### combo
- dose 0.5 all moved;
- dose 1.5 2 subjects 2 did not move
- dose 2,2.5,3 9/10 did not flick their tail

*what is a heat stimulus?*
```{r, message=FALSE, warning=FALSE}
# Using ggplot to creat various boxplots with various features
morphine.box <- ggplot(data = pain, aes(x="",y= Morphine)) + 
    geom_boxplot(fill="lightblue", alpha=0.2, outlier.shape = 21, outlier.color = "black") +
    stat_boxplot(geom='errorbar', linetype=1, width=0.25) + 
    ylab("Morphine") + # Label names
    theme_classic() # A classic theme, no grid lines

marijuana.box <- ggplot(data = pain, aes(x="",y= Marijuana)) + 
    geom_boxplot(fill="lightblue", alpha=0.2, outlier.shape = 21, outlier.color = "black") + 
    stat_boxplot(geom='errorbar', linetype=1, width=0.25) + 
    ylab("Marijuana") + # Label names
    theme_classic() # A classic theme, no grid lines

flick.box <- ggplot(data = pain, aes(x="",y= Flick)) + 
    geom_boxplot(fill="lightblue", alpha=0.2, outlier.shape = 21, outlier.color = "black") + 
    stat_boxplot(geom='errorbar', linetype=1, width=0.25) + 
    ylab("Flick") + # Label names
    theme_classic() # A classic theme, no grid lines

prop.box <- ggplot(data = pain, aes(x="",y= Prop)) + 
    geom_boxplot(fill="lightblue", alpha=0.2, outlier.shape = 21, outlier.color = "black") + 
    stat_boxplot(geom='errorbar', linetype=1, width=0.25) + 
    ylab("Prop") + # Label names
    theme_classic() # A classic theme, no grid lines

# Places the graphs into one space
plot_grid(morphine.box, marijuana.box, flick.box, prop.box, labels = "AUTO", 
  label_fontfamily = "serif",
  label_fontface = "plain",
  label_colour = "lightblue")
```

*maybe 3 boxplots with prop as y, and x as the doses*
**Lucia comment:** not sure I understand this haha (self-reminder to ask)

```{r}
plota<- pain1 %>% filter(Set=='A')%>% ggplot(aes(x=Morphine, y=Prop))+
  geom_boxplot(fill="lightblue", alpha=0.2, outlier.shape = 21, outlier.color = "black") +
    stat_boxplot(geom='errorbar', linetype=1, width=0.25) + 
  xlab("Morphine")+
    theme_classic()
plotb<-pain1 %>% filter(Set=='B')%>% ggplot(aes(x=Marijuana, y=Prop))+
  geom_boxplot(fill="lightblue", alpha=0.2, outlier.shape = 21, outlier.color = "black") +
    stat_boxplot(geom='errorbar', linetype=1, width=0.25) + 
  xlab("Marijuana")+
    theme_classic()
plotab<-pain1 %>% filter(Set=='AB')%>% ggplot(aes(x=Morphine, y=Prop))+
  geom_boxplot(fill="lightblue", alpha=0.2, outlier.shape = 21, outlier.color = "black") +
    stat_boxplot(geom='errorbar', linetype=1, width=0.25) + 
  xlab("Both drugs")+
    theme_classic()# remember they are given the same amount of both drugs, decide whether to change the x axis

plot_grid(plota, plotb, plotab) 
  
```


```{r, message=FALSE, warning=FALSE}
# Used ggpairs to create a scatterplot matrix, did not include Sets becuase of redundant informtion
scaterrplot <- ggpairs(pain, title = "Scatterplot Matrix",
                      columns = c(1:4),
                      proportions = "auto",
                      columnLabels = c("Morphine","Marijuana","Flick",
                      "Prop"),
                      upper = list(continuous = wrap('cor',size = 3)),) 

scaterrplot
```

```{r}
# Create heatmap to show variable correlation
# Round the correlation coefficient to two decimal places
cormat <- round(cor(pain[, -5]), 2)

# Use correlation between variables as distance
reorder_cormat <- function(cormat){ 
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
return(cormat)
}

# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)

# Keeping only upper triangular matrix
# upper_tri returns TRUE/FALSE for each coordinate (TRUE -> part of upper triangle)
# multiplying will thus keep the upper triangle values and set the others to 0
cormat <- cormat*upper.tri(cormat, diag = TRUE)
# Values of the lower triangle (0) are replaced by NA
cormat[cormat == 0] <- NA

# Melt the correlation matrix
cormat <- reshape2::melt(cormat, na.rm = TRUE)

# Create a ggheatmap with multiple characteristics 
ggplot(cormat, aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "purple", high = "blue", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab", name="Pearson\nCorrelation") +
  ggtitle("Correlation Heatmap") + # Title name
  theme_minimal() + # Minimal theme, keeps in the lines
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) +
  coord_fixed() +
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 2)
```
* bar chart for the responsivity to the drugs
* "dots" chart in 4 different steps

### Synergy between the 2 drugs
*idek what he means with this we should ask*
yes 
### At what dosage 50% of the subjects are expected to respond?




